---
title: "lidR NEON example"
author: "Tyson Lee Swetnam"
date: "4/11/2018"
output: html_document
---

# Individual plant / tree segmentation with lidR

[lidR](https://github.com/Jean-Romain/lidR/wiki) is a R package created by Jean-Romain. It is built natively in R, and allows users to select from multiple segmentation techniques (e.g. watershed, 

Example scripts provide below were written by Andrew Sanchez-Meador (Northern Arizona University), adapted for this example by Tyson L. Swetnam (University of Arizona).

```{r, echo=FALSE}
# Check for packages
packageVersion('IRdisplay')
packageVersion('repr')

# Working in a Jupyter Notebook?
# packageVersion('IRkernel')

# install.packages("rjson")
# install.packages("rgdal")
# install.packages("lidR")
# install.packages("IRdisplay")
```

```{r, echo=FALSE}
library(lidR)
library(rgdal)

library(dplyr)
library(ggplot2)
library(rjson)
library(bitops)
library(jsonlite)
library(leaflet)
library(RCurl)
```

```{r}
lidr_options(verbose = TRUE, progress = TRUE)
```
## Build a project 

Set the working directory where the lidar tiles are, and where the output will be

```{r}
getwd()
```

```{r}
setwd('/qubes/lessons/srer_laz/classified')
```

Create a catalog of the entire flight (at the directory level)

```{r}
project <- catalog("/qubes/lessons/srer_laz/classified//NEON_D14_SRER_DP1_514000_3519000_classified_point_cloud.laz")
plot(project)                   
```

Read file information

```{r}
las = readLAS("/qubes/lessons/srer_laz/classified//NEON_D14_SRER_DP1_514000_3519000_classified_point_cloud.laz")
las
```
## Set some global catalog options

For this dummy example, the clustering size is 500 m and the buffer is 15 m using
a single core (because this example is run on the CRAN server when the package is submitted).

```{r}
catalog_options(buffer = 15, multicore = 16, tiling_size = 500)
```

## Build the function that analyzes each cluster of the catalog.

The function's first argument is a LAS object. The internal routine takes care of this part. The other arguments can be freely choosen by the user. See the following template:

```{r}
tree_area = function(las)
{
  if (is.null(las))
    return(NULL)
  
  # segment trees (in this example the low point density does not enable
  # accurate segmentation of trees. This is just a proof-of-concept)
  lastrees(las, algorithm = "li2012")
  
  # Here we used the function tree_metric to compute some metrics for each tree. This
  # function is defined later in the global environment.
  m = tree_metrics(las, myMetrics(X, Y, Z, buffer, treeID))
  
  # If min buffer is 0 it means the trees were at least partly in the non-buffered area, so we
  # want to keep these trees.
  # However, the trees that are on the edge of the buffered area will be counted
  # twice. So we must remove the trees on the right side and on the top side of the buffer
  # If max buffer is <= 2 it means that the trees belong inside the area of interest, on
  # the left side or the bottom side, or both.
  m = m[minbuff == 0 & maxbuff <= 2]
  
  # Remove buffering information that is no longer useful
  m[, c("minbuff","maxbuff") := NULL]
  
  return(m)
}
```

This function enables users to extract, for a single tree, the position of the highest point and some information about the buffering position of the tree. The function tree_metrics takes care of mapping along each tree.

```{r}
myMetrics <- function(x, y, z, buff, treeID)
{
  i = which.max(z)
  xcenter = x[i]
  ycenter = y[i]
  A = area(x,y)
  dat = matrix(c(x, y), nrow=length(x), ncol=2)
  ch = chull(dat)
  coords <- dat[c(ch, ch[1]), ]  
  sp_poly <- SpatialPolygons(list(Polygons(list(Polygon(coords)), ID=treeID[i])))
  minbuff = min(buff)
  maxbuff = max(buff)
  
  return(
    list(
      x = xcenter,
      y = ycenter,
      area = A,
      crowns=list(sp_poly),
      minbuff = minbuff,
      maxbuff = maxbuff
    ))
}
```

Everything is now well defined, so now we can process over an entire catalog with hundreds of files.

## Process the project. 

The arguments of the user-defined function must belong in a labelled list. We also pass extra arguments to the function readLAS  to load only X, Y and Z coordinates. This way we save a huge amount of memory, which can be used for the current process.

```{r}
output = catalog_apply(project, tree_area, select = "xyz")
output = data.table::rbindlist(output)
list_of_SPols = output$crowns
```

Now remove the crown information from the output, to prepare to write the geojson.

```{r}
output = output[,1:4]
```

Set the spatial coordinates and projection for the treelocs output file and write the geojson

```{r}
coordinates(output)<-~x+y
proj4string(output) <- CRS("+proj=utm +zone=12 +datum=WGS84")
writeOGR(obj=output, dsn="/qubes/lessons/srer_laz/treelocs.geojson", layer="treeslocs", driver="GeoJSON", overwrite_layer = T)
```

Now we need to manipulate the crown polys to be able to coerse them to a SpatialPolygonsDataFrame

```{r}
Spols <- SpatialPolygons(lapply(1:length(list_of_SPols), function(i) {
  Pol <- slot(list_of_SPols[[i]], "polygons")[[1]]
  slot(Pol, "ID") <- as.character(i)
  Pol
}))
```

Get the polygon IDs

```{r}
pid <- sapply(slot(Spols, "polygons"), function(x) slot(x, "ID"))
```

Create dataframe with correct rownames

```{r}
p.df <- data.frame( ID=1:length(Spols), row.names = pid)
```

coersion to SpatialPolygonsDataFrame and write it out

```{R}
Spols <- SpatialPolygonsDataFrame(Spols, p.df)
writeOGR(obj=Spols, dsn="/qubes/lessons/srer_laz/trees.geojson", layer="trees", driver="GeoJSON", overwrite_layer = T)
```

Plot it for fun

```{r}
library("leaflet")

trees <- geojsonio::geojson_read("/qubes/lessons/srer_laz/trees.geojson")

epsg32612 <- leafletCRS(crsClass = "L.Proj.CRS", code = "EPSG:32612", 
    proj4def = "+proj=utm +zone=12 +datum=WGS84 +units=m +no_defs", 
    resolutions = 2^(13:-1), 
    origin = c(0, 0)
)

tile_url <- "http://api.geosition.com/tile/osm-bright-3006/{z}/{x}/{y}.png"
tile_attrib <- "Map data &copy; <a href='http://www.openstreetmap.org/copyright'>OpenStreetMap contributors</a>, Imagery &copy; 2013 <a href='http://www.kartena.se/'>Kartena</a>"

leaflet(options = leafletOptions(worldCopyJump = F, crs = epsg32612)) %>%
  setView(11.965053, 57.70451, 13) %>%
  addTiles(urlTemplate = tile_url,
    attribution = tile_attrib,
    options = tileOptions(minZoom = 0, maxZoom = 14, continuousWorld = T)) %>%
  addMarkers(11.965053, 57.70451)
```

